Toward Supervised Reinforcement Learning with Partial States for Social HRI

### 2017-08-28 paper.tex
Update on Overleaf.

    Social interacting is a complex task for which machine learning holds
    particular promise. However, as no sufficiently accurate simulator of human
    interactions exists today, the learning of social interaction strategies has
    to happen online in the real world. Actions executed by the robot impact on
    humans, and as such have to be carefully selected, making it impossible to
    rely on random exploration. Additionally, no clear reward function exists
    for social interactions. This implies that traditional approaches used for
    Reinforcement Learning cannot be directly applied for learning how to
    interact with the social world. As such we argue that robots will profit
    from human expertise and guidance to learn social interactions. And as the
    quantity of input a human can provide is limited, new methods have to be
    designed to use human input more efficiently. In this paper we describe a
    setup in which we combine a framework called Supervised Progressively
    Autonomous Robot Competencies (SPARC), which allows safer online learning
    with Reinforcement Learning, with the use partial states rather than full
    states to accelerate generalisation and obtain more quickly a usable action
    policy.

@@

In previous work~\cite{senft2015sparc,senft2017supervised}, we introduced the
Supervised Progressively Autonomous Robot Competencies (SPARC), a way to teach a
robot an action policy while interacting based on a human supervisor intervening
and correcting actions before they were executed by the robot. In this paper, we
propose to extend this approach to allow the supervisor to highlight features in
the environment relevant for the selected action. During the action selection
phase, the robot can compare these features, defined as partial states, with the
current state to select an action. The selected action is presented to the human
supervisor, who can either correct the action or approve it for execution.

@@

Initial knowledge can be provided to the agent through demonstrations. These
demonstrations can be used to create an initial policy which is sufficiently
efficient to start interacting in the environment and gather information to
improve over time. This initial policy, required to receive meaningful feedback
from the environment can be impossible to reach by relying only on random
exploration and feedback from the environment. For example the game of Go in its
larger board size contains more than $10^{210}$ states, so exhaustive search is
not possible.  Silver et al.~\cite{silver2016mastering} started with supervised
learning from Go masters' games to learn a decent enough policy and then
proceeded using deep learning, self play and tree search to achieve super-human
capabilities and beating the best human players.

@@

Another approach to guide the learning is to bias the action selection.
In~\cite{thomaz2008teachable}, the authors propose to use a human supervisor to
supply an agent with both rewards and potential guidance indicating what the
agent should pay attention to, or what action it should take next. They show
that giving the power to the supervisor to bias the action selection can improve
the learning, making it faster and safer.

In~\cite{senft2015sparc}, we introduced the Supervised Progressively Autonomous
Robot Competencies (SPARC). SPARC relies on a supervisor with the ability to
control the robot's actions. This supervisor is presented with the action the
robot is about to execute. He can then choose to cancel it, allow it to be
executed, or can select an alternative action. Based on the supervisor's
decisions, the robot learns which actions are desired and can as such refine its
policy over time reducing the need for the supervisor to correct or select an
action.

SPARC gives control of the robot's actions to an expert, who can guide the
exploration in the desired part of the environment which ensures the robot's
behaviour is consistently appropriate. As the exploration is guided, and all the
actions are useful, the learning can be faster than autonomous learning or
learning based on human feedback as illustrated in Figure~\ref{fig:comparison}.
In Autonomous Learning, the agent has first to discover its environment to start
gathering relevant feedbacks, leading to a low initial performance and a slow
improvement in early stage of learning. With a teaching based on human
feedbacks, the teacher can quickly provide feedbacks on actions to reach more
quickly an efficient policy. However, as the teacher only provide feedbacks and
cannot prevent the agent to make mistakes, the initial performance can be poor.
On the other hand, SPARC, with the control of the teacher over the executed
actions can prevent the agent to make mistakes in early stages, achieving a high
performance even at the start of the learning.

@@

\begin{figure}
    \centering
    \includegraphics[width=0.9\linewidth]{./figs/motivation.pdf}
    \caption{An illustration of the evolution over time of the performance,
    autonomy and human workload for an autonomous learner, an approach using
    human feedbacks, and SPARC.}
    \label{fig:comparison}
\end{figure}

In~\cite{senft2017supervised}, we presented a way to combine SPARC and RL, by
assigning a positive reward to every action executed by the robot, making the
assumption that every action has been explicitly or implicitely approved by the
supervisor. However this method was directly mapping a single (state, action)
pair to a reward without making use of any kind of generalisation. As such it
was not applicable to environments with a continuous or high dimensional space
or non-deterministic transitions from one state to another, elements which are
typically present in social interactions. Similarly to TAMER or COACH, this
approach allows to reproduce an action policy from a teacher even in the absence
of a reward function, but as the teacher has control over the robot actions, the
learning can be faster and safer.

@@

To make RL applicable in high dimensional or continuous states, the algorithm
requires a way to generalise knowledge to unseen states. A classic approach is
to use a feed-forward neural network or deep learning to learn a the value
function generalising to unseen states. Neural network rely on having a large
number of datapoints to converge toward a good function approximator.
Alternatively, Cobo et al. proposed in~\cite{cobo2011automatic} to abstract
features automaticly from demonstrations to learn faster. But even in this case,
the number of datapoints required is still high (around 1000 samples per
participants). However, in many applications and especially in HRI, these
amounts of data are not available or obtainable due to practical constraints.
Furthermore, human responses are often noisy and lack consistency, so methods
are needed which can generalise from a low number of noisy data points.

In this paper we propose to use a human user to highlight the relevant features
of the environment to reduce the state dimension of the points only to relevant
information. We introduce the concept of a partial state, a sliced version of
the state defined only on a subset of the dimensions of the state. This shifts
the (state, action) pair paradigm to (partial state, action)  and in the case of
RL, the tuple (state, action, reward) to (partial state, action, reward). This
allows to compare the current state and the datapoints only on relevant features
for action selection. With this instance-based method~\cite{aha1991instance},
the algorithm can have a state abstraction allowing it to generalise even with a
few datapoints instead of the large numbers normally required to abstract
features from examples.

@@

The GUI (Figure~\ref{fig:gui}) is an augmented version of the game itself, which
can be use to make the robot move items on the game by dragging them on the GUI
or which can display actions proposed by the robot with a cancel button to
refuse an action. For example in Figure~\ref{fig:gui} the robot proposes to move
the eagle to the rat, and highlight (as shown by blue circles) the eagle and the
rat. This indicate that features relevant to the eagle and the rat have been
used to select this action.

@@

\begin{figure}
        \centering
    \includegraphics[width=60mm]{./figs/proposition.png}
    \caption{Interface for the supervisor with an action being proposed, moving
        the eagle to the rat with highlight of the eagle and the rat.}
        \label{fig:gui}
\end{figure}

@@

The current work could also be extended to allow the agent to continue to
improve its behaviour even in the absence of a supervisor, progressively
exploring around the learnt policy improving its behaviour beyond the
demonstrations. This could be done by allowing the supervisor to provide rewards
during the learning phase and combine these rewards with the demonstrations to
learn a reward function in a fashion similar to Inverse Reinforcement
Learning~\cite{abbeel2004apprenticeship} or TAMER ~\cite{knox2009interactively}.
This could also use partial states associated to these rewards to ease the
generalisation of the reward function with only a low number of datapoints.  But
without the supervisor, the assumption that only a myopic action selection is
sufficient would not hold anymore and the problem of delayed rewards would have
to be tackled.

>>>>

    Social interacting is a complex task for which machine learning holds
    particular promise. However, as no sufficiently accurate simulator of human
    interactions exists today, the learning of social interaction strategies has
    to happen online in the real world. Actions executed by the robot impact on
    humans, and as such have to be carefully selected, making it impossible to
    rely on random exploration. Additionally, no clear reward function exists
    for social interactions. This implies that traditional approaches used for
    Reinforcement Learning cannot be directly applied for learning how to
    interact with the social world. As such we argue that robots will profit
    from human expertise and guidance to learn social interactions. However, as the
    quantity of input a human can provide is limited, new methods have to be
    designed to use human input more efficiently. In this paper we describe a
    setup in which we combine a framework called Supervised Progressively
    Autonomous Robot Competencies (SPARC), which allows safer online learning
    with Reinforcement Learning, with the use of partial states rather than full
    states to accelerate generalisation and obtain a usable action
    policy more quickly.

@@

In previous work~\cite{senft2015sparc,senft2017supervised}, we introduced the
Supervised Progressively Autonomous Robot Competencies (SPARC) as a way to teach a
robot an action policy while interacting based on a human supervisor intervening
and correcting actions before they were executed by the robot. In this paper, we
propose to extend this approach to allow the supervisor to highlight features in
the environment relevant for the selected action. During the action selection
phase, the robot can compare these features, defined as \textit{partial states}, with the
current state to select an action. The selected action is presented to the human
supervisor, who can either correct the proposed action or approve it for execution.

@@

Initial knowledge can be provided to the agent through demonstrations. These
demonstrations can be used to create an initial policy which is sufficiently
efficient to start interacting in the environment and gather information to
improve over time. This initial policy, required to receive meaningful feedback
from the environment can be impossible to reach by relying only on random
exploration and feedback from the environment. For example the game of Go in its
larger board size contains more than $10^{210}$ states, so exhaustive search is
not possible.  Silver et al.~\cite{silver2016mastering} started with supervised
learning from Go masters' games to learn a decent enough policy and then
proceeded using deep learning, self play and tree search to achieve super-human
capabilities and the capacity to beat the best human players.

@@

Another approach to guide the learning is to bias the action selection.
In~\cite{thomaz2008teachable}, the authors propose using a human supervisor to
supply an agent with both rewards and potential guidance indicating what the
agent should pay attention to, or what action it should take next. They show
that giving the power to the supervisor to bias the action selection can improve
the learning, making it faster and safer.

In~\cite{senft2015sparc}, we introduced the Supervised Progressively Autonomous
Robot Competencies (SPARC). SPARC relies on a supervisor with the ability to
control the robot's actions. This supervisor is presented with the action the
robot is about to execute. He can then choose to cancel it, allow it to be
executed, or can select an alternative action. Based on the supervisor's
decisions, the robot learns which actions are desired and can as such refine its
policy over time, thus progressively reducing the need for the supervisor to correct or select an
action.

SPARC gives control of the robot's actions to an expert, who can guide the
exploration in the desired part of the environment which ensures the robot's
behaviour is consistently appropriate. As the exploration is guided, and all the
actions are useful, the learning can be faster than autonomous learning or
learning based on human feedback as illustrated in Figure~\ref{fig:comparison}.
In Autonomous Learning, the agent has first to discover its environment to start
gathering relevant feedback, leading to a low initial performance and a slow
improvement in early stage of learning. With teaching based on human
feedback, the teacher can quickly provide information on actions to reach an efficient policy more
quickly. However, as the teacher only provides feedback and
cannot prevent the agent making mistakes, the initial performance can be poor.
On the other hand, SPARC, with the control of the teacher over the executed
actions, can prevent the agent from making mistakes in early stages, achieving a high
performance even at the start of the learning.

@@

\begin{figure}
    \centering
    \includegraphics[width=0.9\linewidth]{./figs/motivation.pdf}
    \caption{An illustration of the evolution over time of the performance,
    autonomy and human workload for an autonomous learner, an approach using
    human feedback, and SPARC.}
    \label{fig:comparison}
\end{figure}
%PEB: in the figure "feedbacks" should be "feedback"

In~\cite{senft2017supervised}, we presented a way to combine SPARC and RL, by
assigning a positive reward to every action executed by the robot, making the
assumption that every action has been explicitly or implicitely approved by the
supervisor. However this method directly mapped a single (state, action)
pair to a reward without making use of any kind of generalisation. As such it
was not applicable to environments with a continuous or high dimensional space
or non-deterministic transitions from one state to another, elements which are
typically present in social interactions. Similarly to TAMER or COACH, this
approach allows to reproduce an action policy from a teacher even in the absence
of a reward function, but as the teacher has control over the robot actions, the
learning can be faster and safer.

@@

To make RL applicable in high dimensional or continuous states, the algorithm
requires a way to generalise knowledge to unseen states. A classic approach is
to use a feed-forward neural network or deep learning to learn a the value
function generalising to unseen states. Neural networks rely on having a large
number of datapoints to converge toward a good function approximator.
Alternatively, Cobo et al. proposed in~\cite{cobo2011automatic} to abstract
features automatically from demonstrations to learn faster. But even in this case,
the number of datapoints required is still high (around 1000 samples per
participant). However, in many applications and especially in HRI, these
amounts of data are not available or obtainable due to practical constraints.
Furthermore, human responses are often noisy and lack consistency, so methods
are needed which can generalise from a low number of noisy data points.

In this paper we propose to use a human user to highlight the relevant features
of the environment to reduce the state dimension of the points only to relevant
information. We introduce the concept of a \textit{partial state}, a sliced version of
the state defined only on a subset of the dimensions of the state. This shifts
the (state, action) pair paradigm to (partial state, action)  and in the case of
RL, the tuple (state, action, reward) to (partial state, action, reward). This
allows a comparison of the current state and the datapoints only on relevant features
for action selection. With this instance-based method~\cite{aha1991instance},
the algorithm can have a state abstraction allowing it to generalise even with a
few datapoints instead of the large numbers normally required to abstract
features from examples.
%PEB: sounds very much like the "don't care" token in learning classifier systems...

@@

The GUI (Figure~\ref{fig:gui}) is an augmented version of the game itself, which
can be use to make the robot move items on the game by dragging them on the GUI
or which can display actions proposed by the robot with a cancel button to
refuse an action. For example in Figure~\ref{fig:gui} the robot proposes to move
the eagle to the rat, and highlights (as shown by blue circles) the eagle and the
rat. This indicate that features relevant to the eagle and the rat have been
used to select this action.

@@

\begin{figure}
        \centering
    \includegraphics[width=60mm]{./figs/proposition.png}
    \caption{Interface for the supervisor with an action being proposed, moving
        the eagle to the rat highlighting both the eagle and the rat.}
        \label{fig:gui}
\end{figure}

@@

The current work could also be extended to allow the agent to continue to
improve its behaviour even in the absence of a supervisor, progressively
exploring around the learnt policy improving its behaviour beyond the
demonstrations. This could be done by allowing the supervisor to provide rewards
during the learning phase and combine these rewards with the demonstrations to
learn a reward function in a fashion similar to Inverse Reinforcement
Learning~\cite{abbeel2004apprenticeship} or TAMER ~\cite{knox2009interactively}.
This could also use partial states associated to these rewards to ease the
generalisation of the reward function with only a low number of datapoints. However,
without the supervisor, the assumption that only a myopic action selection is
sufficient would not hold anymore and the problem of delayed rewards would have
to be tackled.
